### 2008-04-30 trunk/textos/dsrt/engaoi.tex
eng

\documentclass[conference]{IEEEtran}

\usepackage{graphicx}
%\usepackage{listings}
\usepackage{algorithm}
\usepackage{algorithmic}
%\usepackage{algorithmicx}
\usepackage{amsmath}

%\usepackage[brazil]{babel}   
\usepackage[latin1]{inputenc}

%\DeclareGraphicsExtensions{eps}

%\newcommand{\malg}{A$^3$}
\newcommand{\malg}{A$^3$}
\newcommand{\cridis}{critical\_distance}
\newcommand{\viewdis}{view\_distance}
\newcommand{\circleaoi}{C}
\newcommand{\circleat}{C \& A}
\newcommand{\fov}{FoV}
\newcommand{\noaoi}{None}

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{A novell interest management algorithm for distributed simulations of MMGs}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
% \author{\IEEEauthorblockN{Carlos Eduardo Benevides Bezerra}
% \IEEEauthorblockA{Federal University of Rio Grande do Sul\\
% Bento Gonçalves, 9500, Porto Alegre, RS, Brazil\\
% E-mail: carlos.bezerra@inf.ufrgs.br}
% \and
% \IEEEauthorblockN{Fábio Reis Cecin}
% \IEEEauthorblockA{Federal University of Rio Grande do Sul\\
% Bento Gonçalves, 9500, Porto Alegre, RS, Brazil\\
% E-mail: frcecin@inf.ufrgs.br}
% \and
% \IEEEauthorblockN{Cláudio Fernando Resin Geyer}
% \IEEEauthorblockA{Federal University of Rio Grande do Sul\\
% Bento Gonçalves, 9500, Porto Alegre, RS, Brazil\\
% E-mail: geyer@inf.ufrgs.br}
% }

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% for over three affiliations, or if they all won't fit within the width
% of the page, use this alternative format:
% 
\author{\IEEEauthorblockN{Carlos Eduardo B. Bezerra,
	Fábio R. Cecin,
	Cláudio F. R. Geyer}
\IEEEauthorblockA{Federal University of Rio Grande do Sul\\
	Bento Gonçalves, 9500, Porto Alegre, RS, Brazil,\\
	Email: \{carlos.bezerra, fcecin, geyer\}@inf.ufrgs.br}}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

\begin{abstract}
%\boldmath
Traditionally, a central server is utilized to provide support to MMGs (massively multiplayer games), where the number of participants is in the order of tens of thousands. Much work has been done trying to create a fully peer-to-peer model to support this kind of application, in order to minimize the maintenance cost of its infrastructure, but critical questions remain. Examples of the problems relative to peer-to-peer MMG support systems are: vulnerability to cheating, overload of the upload links of the peers and difficulty to maintain consistency of the simulation among the participants. In this work, it is proposed the utilization of geographically distributed lower-cost nodes, working as a distributed server to the game. The distribution model and some related works are also presented. To address the communication cost imposed to the servers, we specify a novell refinement to the area of interest technique, significantly reducing the necessary bandwidth. Simulations have been made with ns-2, comparing different area of interest algorithms. The results show that our approach achieves the least bandwidth utilization, with a 33.10\% maximum traffic reduction and 33.58\% average traffic reduction, when compared to other area of interest algorithms.

\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the conference you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals/conferences frown on
% math in the abstract anyway.

% no keywords

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
% no \IEEEPARstart
In the last years, electronic games have become very popular, specially massively multiplayer games, where the number of simultaneous players is in the order tens of thousands \cite{cecin2004fsa}. Worlf of Warcraft \cite{worldofwarcraft}, Lineage II \cite{lineage2} and Guild Wars \cite{guildwars} are successful examples.

Usually, the network support for this kind of application consists in a central server with plenty of resources (processing power and available bandwidth), which accepts connections from the clients. Each player interacts through one of these clients, which sends his actions to the server, that processes them, calculating their influence over the game, and broadcasts the action's result to every client. Due to the number of simultaneous participants that this kind of game usually has, these tasks demand a significant amount of resources from the server, which receives and processes the actions from all the players and broadcast state updates to all of them.

Recently, some alternatives to the central server approach have been researched. One of them is to distribute among the own participants the game simulation and the task to broadcast the state updates when they perform actions. Communication between them is in a peer-to-peer manner, forming a decentralized network \cite{schiele2007rpp}. This approach would be the ideal if it did not have not some critical inherent drawbacks. For example, as the player computers participate on the simulation, they need to agree on the resulting state of the game match. If this agreement does not happen, some inconsistency on the game state among the players may occur.

There is another question, concerning the number of trasmissions that each peer needs to execute. In the client-server model, each client only needs to send his actions to the server, which processes them and broadcasts the resulting game state to every player. In the peer-to-peer model, each peer becomes responsible for processing its own player's actions and send the resulting state to everyone else to whom that state change might be relevant. The problem is that we cannot assume that every client has enough bandwidth to do this. Besides, without a central server, the game becomes dependant of the peers' simulation, which can be corrupted in order to result in an invalid state which benefits a player, or even invalidate the whole game session.

Besides the peer-to-peer model, another alternative is to use a distributed server, in which nodes connected between them divide the game simulation and the state update broadcasts \cite{assiotis2006dam}. Such approach allows for the use of lower-cost computers to form the distributed server system, reducing the cost of the support infrastructure. Also, consistency maintenance becomes easier than with the peer-to-peer model, because a single server decides the simulation for its assigned portion of the world, and problems related to malicious players may be abstracted, since the servers are able to verify the simulation and detect cheating. Finally, with less bandwidth and processing power requirements for the clients, the game becomes accessible for a wider public.

%Alguns trabalhos [citações] têm proposto técnicas para reduzir a carga de processamento e comunicação associada a servidores de jogos maciçamente multijogador, sendo uma alternativa a distribuição desta carga entre diversos servidores, cada um com custo de aquisição e manutenção menor do que o de um grande servidor central. Ainda assim, com as atuais técnicas de distribuição e modelo de comunicação entre os servidores e os clientes, têm-se um alto custo para manter o sistema servidor distribuído como um todo.
However, to avoid the distributed server system maintenance cost to be the same as the central server's, it is necessary to perform some optimizations, in order to reduce the necessary bandwidth for each server node. The present work proposes an algorithm to reduce the bandwidth usage caused by the game traffic between servers and clients, through a refinement of the players' interest management technique \cite{boulanger2006cim}. The basic principle of this technique is that each participant of the game receives only state updates that are relevant to them. Simulations have been made, comparing our proposed algorithm with conventional ones, obtaining significant results.

This paper is organized as follows: in section \ref{relatedwork}, some related works with proposed distribution models are presented; in section \ref{def}, some necessary definitions are made; in section \ref{model}, we propose a distribution model to be used as basis of our approach; in section \ref{aoi}, we review some existing interest management algorithms; in section \ref{gaoi}, we present our approach in detail; in sections \ref{sim} and \ref{results}, the simulation and its results, respectively, are described and in section \ref{conclusion}, we present our final considerations.

\section{Related Work}
\label{relatedwork}
%talvez deva reduzir esta seção.

In the past few years, many works have been done trying to distribute the massively multiplayer games support. One approach is the peer-to-peer model, which has some issues concerning game state consistency among the the participating peers, vulnerability to cheating, and increased bandwidth usage by each peer. There are some proposals that try to minimize these problems. One of then is presented in \cite{schiele2007rpp}, where is suggested the division of the virtal simulated game environment in regions, and inside each region, a peer is chosen to be the region coordinator. Its function will be manage the players' interest, checking to what peers each update really needs to be sent. In this way, the peers' upload bandwidth usage is reduced. However, this usage will be still greater than with a server to perform the broadcasts. In the peer-to-peer model, each player usually must send updates to more than one other player. Besides, the chosen peer must be trustable to be the peer coordinator.

%figura

Another work focused on the peer-to-peer model \cite{iimura2004zfg} uses an approach similar to the one presented in \cite{schiele2007rpp}, but suggests, for each region of the virtual environment, the creation a ``zoned federation'' formed by peers chosen among the participants in that region. Since different nodes manage that region, and will need to agree in order to proceed the game, the simulation becomes more trustable. However, the risk of the chosen peers commit collusion cheating \cite{yan2005scc} is not eliminated. Besides, the agreement itself between the peers - that provides increased reliability - creates a fair amount of extra traffic between the participants of the federation, possibly delaying each simulation step.

%figura

A key problem in the peer-to-peer architectures, in what concerns the use of interest management, is that some peers are responsible for part of the simulation and for decide to whom each state update would interest. Assuming that there are only trustable players, the interest management technique may be useful. However, supposing that a player is malicious, he could hack his client software and avoid sending state updates to certain players, who would be in disadvantage. Consequently, the distributed server model is considered more suitable to use the interest management technique.

An example of this kind of model is described in \cite{assiotis2006dam}, where a distributed architecture for MMGs is proposed. It is also based on the division of the virtual environment in regions, each one with a server node assigned to it. The player whose avatar is situated in a region of the game world should connect to the server responsible for that region. This way, each server would group different players, based on their locality on the virtual space. To achieve consistency between the different server nodes performing the simulation, it is used the concept of locks. When a server node needs to alter the state of some entity of the game, it first needs to obtain exclusive access to that entity. To do that, it negociates with the other server nodes which might also need to make some change in the same object, and then obtains the lock. When it finishes changing the state, it releases the lock and notifies the other servers

%figura

The first major restriction in the proposal of \cite{assiotis2006dam}, however, is the assumption that all server nodes are connected with one another through a high velocity and low latency network, what cannot be assured when using lower-cost geographically distributed nodes. Another problem is that the scalability question is addressed simply via the expansion of the virtual environment area, supposing that the players will spread over it. Finally, they suggest to solve the hotspot problem through successive recursive environment partitioning, until the number of players per server is below a certain threshold. However, there is a practical limit to the repartitioning of the virtual environment, and they do not suggest what to do when this limit is reached.

%falo de ng?

\section{Definitions}
\label{def}

%Será utilizado o termo cliente para referir-se ao computador utilizado por cada jogador para conectar-se a um dos servidores do jogo, assim como o termo servidor fará referência a cada nodo integrante do sistema distribuído que estará servindo o jogo.
It is necessary to describe the network support model on which the proposed interest management algorithm is intended to be utilized. Throughout the text, some terms will be used, and their definitions are given here:

%É necessário descrever o modelo de suporte a jogos maciçamente jogador sobre o qual pretende-se utilizar o algoritmo de gerenciamente de interesse proposto. Ao longo do texto, serão utilizados alguns termos que precisam antes ser definidos: %Define-se avatar como sendo a representação no ambiente virtual do jogador, que, através dele, interage com o mundo do jogo e com outros jogadores. Regiões são divisões do ambiente virtual, cada uma podendo conter jogadores presentes. Assumindo regiões contíguas, define-se fronteira como sendo a divisa entre duas regiões adjacentes.

\textbf{Avatar} is the player representation in the virtual environment. Through his avatar, the player interacts with the game world and with the other players. Examples of avatars are the characters in MMORPG (massively multiplayer online role-playing games) games like World of Warcraft.

\textbf{Entities} are the constituent parts of the virtual world. Examples of entities are the avatars of the players as well as avatars controlled by artificial intelligence of the server - monsters of MMORPGs, for example - and objects in the environment, such as doors, weapons and items with which avatars can interact.

\textbf{State} is the set of properties that can be observed in the various entities of the game. The overall state of the simulated world is composed of individual states of the different entities present in it.

The players interact with the game world through \textbf{actions}. An action is a command of the player as, for example, move your avatar for particular location in the virtual world, attacking another player, take some object available in the environment and so on. In general, actions change the state of one or more entities in the game.

A \textbf{region} is a partition of the virtual environment, under the responsibility of a single server. Thus, players whose avatars are located in the same region will have its interaction improved, since their clients are connected to the same server.

The \textbf{border} between two regions is the line that divides the areas that these regions occupy. When an avatar is located near a border, the server responsible for the region beyond this border is notified about the presence of that avatar by the server where it is.

\section{Distribution Model}
\label{model}

This work is based on a virtual environment partitioned into regions, each managed by a server. The regions are contiguous, exploring the locality of the players' avatars. Thus, avatars next to each other will probably be located in the same region and, therefore, their clients will tend to be connected to the same server, so that their interaction is faster (Fig. \ref{partitions}). Situations in which two players interacting with each other are connected to different servers imply a higher bandwidth usage, as it is required some sort of negotiation between the servers to which the different players are connected, in order to maintain the states of simulation in both servers identical. Also, the hop count of each message between the clients of these players is increased, since there is one more intermediary.

\begin{figure}[!t]
	\centering
	%\includegraphics[width=2.5in]{max}
	\includegraphics[width=0.9\linewidth]{part}
	\caption{Distribution model}
	\label{partitions}
\end{figure}

One issue that relates to that type of partitioning of the virtual environment concerns the borders between regions. If an avatar of a client connected to a server is close to the border of a region with another, which is managed by another server, it will be necessary to have an exchange of information between these servers. This information consists of state updates of the entities that are interacting with each other despite being located in different regions. For example, let $S_A$ be the server responsible for the region $R_A$ where is located the avatar of the client $C_A$ and let be $S_B$ the server responsible for another region, $R_B$, where is located the avatar of the client $C_B$. When the avatar of $C_A$ gets close the border with $R_B$, $S_A$ sends to $S_B$ a message notifying it about the presence of that avatar near its region's border. If the avatar of $C_B$ also gets close to the border with $R_A$ and to the avatar of $C_A$, $S_B$ sends another message, notifying $S_A$. Then, the avatars of $C_A$ and $C_B$ start exchanging state updates through their servers.

Regarding the simulation of the actions performed by players whose avatars are located in different regions, it must be decided how it will be performed. As the focus of this work is not the simulation itself, but the bandwidth usage optimization through a new interest management algorithm, it was decided that the simulation will be performed by the server to which the client of that player is connected. Thus, if the player whose avatar is in region $R_i$ performs an action near the border involving entities in $R_j$, it is the server $S_i$ who will decide the outcome of these actions, forwarding only the already calculated new state to $S_j$.

This way, the details of this mechanism will not result in relevant changes to the interest management. When a player $J$, whose avatar is near the border of a region, performs actions whose results must be broadcasted to players with their avatars in other regions, the server $S$, responsible for $J$, simulates his actions, calculates the resulting state and simply sends it to the neighbor server, as if it was sending to its own clients. Similarly, when $S$ receives the resulting state of the action of a player who is connected to a neighbor server, it broadcasts the state to the clientes connected to it as if a player in its own region had executed that action.

%figura

%Cada servidor terá uma lista dos endereços e portas dos outros servidores, assim como a região a que cada um está associado. Quando um jogador deseja participar do jogo, deve conectar-se a um servidor qualquer. Se o jogador entrar com um avatar novo, é escolhida alguma região para ele começar a jogar, e o servidor a que ele se conectou irá redirecioná-lo para o servidor correto. O critério de escolha pode ser o número de jogadores em cada região, como também pode ser feita simplesmente uma seleção aleatória da região. Se já tiver jogado antes, será redirecionado para o servidor responsável pela área na qual ele parou de jogar da última vez. Desta forma, aumenta-se a probabilidade do jogador recomeçar a partida com as informações mais recentes a respeito de seu estado.

\section{Interest Management Algorithms}
\label{aoi}

%como há vários servidores, o processamento das áreas de interesse é paralelizado, também, então não é tãaao ruim calcular a distância entre cada par de avatares..
%explicar uma por uma, como funciona e o porquê.

In order to provide an identical sense of the environment among the players, each one of them must maintain a copy of the entities' states, which must be the same for everyone. The simplest way to do this is to broadcast the states of all entities to all clients. The problem of this approach is that it generates a heavy traffic between the servers and clients, preventing the game to well, as the number of participants increase. To save bandwidth, both of the players, as the servers that intermediate them, a technique known as interest management is employed. This technique reduces the number of updates that the players will receive - and send, in the case of a peer-to-peer architecture.

In short, the interest management technique works as follows: for every state change of each entity, it is calculated for whom it will be important. For example, if an avatar is miles away from another in the virtual environment, it is most likely that their state changes is irrelevant to each other. Thus, it is not necessary for them to exchange information on their state. This principle - locality - is used as the main criterion in the interest management algorithms.

The algorithms described in the following sections are mainly based on the euclidian distance between each avatar and all other entities in the virtual environment. This could create a scalability problem due to the processing of the distances, but a distributed architecture is assumed, where the processing can and should be parallelized. In the distribution model defined previously, each server controls a region of the map. Therefore, each one of them manages only a subset of the entities of the game, checking only the distances between each pair of them, in addition to the entities that are in a neighbouring region, close to its border.

In the next sections, some versions of this technique, such as circular area based and avatar's field of view based interest management, will be presented. In the section \ref{gaoi}, it is introduced the approach of attenuating the frequency of updates, and a proposed algorithm will be described in detail.

\subsection{Circular area of interest}
\label{circle}

The simplest way to execute interest management is to define a circular area, whose center is determined by the coordinates of the avatar's location in the virtual environment. After that, the euclidian distance between the avatar and every other other entity in the game world is calculated. Let $A$ be an avatar, whose area of interest is a circle of radius $rad$. If an entity $E$ is at a distance less than $rad$ from $A$, then its state updates will be relevant to $A$. $A$ will not receive state updates from entities which are at a distance greater than $rad$. Figure \ref{fcirc} illustrates this type of area of interest.

\begin{figure}[!t]
	\centering
	%\includegraphics[width=2.5in]{max}
	\includegraphics[width=0.9\linewidth]{circle}
	\caption{Circular area of interest}
	\label{fcirc}
\end{figure}

\subsection{Field of view}
\label{angle}

A more refined method to manage the interests of the avatars is to take into account what each player can see. The area within which the player perceives state changes can be defined as a circular sector. This is similar to the circular area of interest described in the previous section, but takes into account that the player can only see objects that are located in front of his avatar.

%pegar isso pra mim!?... CLARO! fui eu que fiz!

One issue to be considered, however, is that the player will not receive state updates from entities entities which are close to his avatar, but behind it. This could cause some problems. For example, if the avatar turns 180 degrees, the player might not be able to see a particular entity that should be there, needing some time to receive its state information. This happens because, despite this entity have been near the avatar, the player did not receive its information yet because it was behind him, outside his field of view. In Figure \ref{fangle} it is illustrated an area of interest that takes into account the player's view angle.

\begin{figure}[!t]
	\centering
	%\includegraphics[width=2.5in]{max}
	\includegraphics[width=0.9\linewidth]{fov}
	\caption{Field of view based area of interest}
	\label{fangle}
\end{figure}

%algoritmo

\section{Graded Area of Interest}
%\subsection{Círculo com atenuação}
\label{gaoi}

The principle behind the approach proposed here is based on the fact that the more distant an entity is from the avatar in the virtual environment, the lower its update frequency for that avatar may be. Therefore, the state updates from entities which are more distant may be received with a longer interval between them. On the other hand, if an entity is very close, it is desirable that the player receives its most recent state change information as soon as possible, to view any alterations quickly.

To achieve this objective, it is necessary to define some parameters:

\textbf{Relevance} - real value ranging from 0 to 1 inclusive, which determines how much an entity's state is relevant to an avatar.

\textbf{Freqüência de atualização} - quantidade de atualizações que cada avatar recebe de cada uma das entidades do ambiente virtual por unidade de tempo.

\textbf{Intervalo normal de atualização} - menor intervalo de tempo entre a chegada de duas atualizações de estado consecutivas de uma mesma entidade em um cliente, ou seja, quando a relevância daquela entidade para o avatar daquele cliente é 1. Assim sendo, o intervalo normal determina a freqüência máxima de atualização.

\textbf{Alcance de visão} - determina a que distância as entidades podem estar do avatar, no máximo, para que o jogador possa visualizá-las.

\textbf{Distância crítica} - é o raio do círculo, em torno do avatar, onde todas as entidades têm relevância igual a 1.

Para se enviar o estado de uma entidade para determinado cliente, verifica-se primeiro quando foi o último envio. O próximo instante de envio é então escalonado para ocorrer após um determinado intervalo de tempo. Se a relevância daquele estado for 1, será utilizado o intervalo normal de atualização. Se for menor que 1, divide-se o intervalo normal pela relevância. Por exemplo, seja um jogo em que o intervalo normal de atualização seja de 200 ms. Se o avatar $A_i$, que acabou de enviar uma atualização de estado para $A_j$, está a uma distância de $A_j$ tal que sua relevância é 0.5, o próximo envio será depois de um intervalo de 200/0.5, ou seja, 400 ms. Apesar deste intervalo ainda ser uma fração de segundo, representa uma diminuição da freqüência de atualização do estado de $A_i$ em 50\%. Como estão a uma distância maior um do outro, e o intervalo foi aumentado de apenas 200 ms, esta variação deverá ser imperceptível para o jogador que controla $A_j$.

É importante perceber que a atenuação da freqüência de atualização das entidades é compatível com outras técnicas mais complexas de gerenciamento de interesse. Em \cite{boulanger2006cim}, são descritos diversos algoritmos de gerenciamento de interesse que poderiam ser ainda melhorados se fosse agregada a idéia de diferentes intervalos de envio baseado na relevância destas atualizações. Geralmente o estado de cada entidade é classificado em um de apenas dois extremos: é relevante ou não é relevante, ignorando-se que há uma vasta gama de valores intermediários. A questão está em como definir esse valor de relevância para cada estado. Nas seções seguintes, serão apresentados dois exemplos de algoritmos que definem o método de se obter esse valor, assim como o tipo de área utilizado. Na seção \ref{csmooth}, é especificado o \malg, que é um algoritmo original de gerenciamento de interesse, que, dentre outros princípios, emprega a atenuação da freqüência de atualização. As simulações e seus resultados são apresentados nas seções \ref{sim} e \ref{results}, respectivamente.

\subsection{Círculo com atenuação}

Um exemplo simples de utilização de intervalos variados de atualização baseado em relevância seria utilizando a área de interesse em formato de círculo. Para obter-se a relevância de uma entidade em relação a um avatar, pode-se fazer com que seu valor seja 1 quando a entidade estiver na mesma posição do avatar e ir diminuindo gradualmente à medida em que se afasta, até chegar a 0, quando para de diminuir não importa o quanto mais se afaste. Essa é uma maneira que, apesar de simples, demonstrou uma significante redução no tráfego entre clientes e servidores. Na figura \ref{graded_circle}, é ilustrado como seria a área de interesse com atenuação gradual da freqüência de atualização das entidades para um determinado avatar.

\begin{figure}[!t]
	\centering
	%\includegraphics[width=2.5in]{max}
	\includegraphics[width=0.9\linewidth]{gcircle}
	\caption{Circular area of interest with update frequency atenuation}
	\label{graded_circle}
\end{figure}

%           A                   A              A 
%\subsection{Ângulo de visão com área próxima e atenuação da freqüência de atualização}
\subsection{Algoritmo \malg}
%View Angle with close area and update frequency atenuation
%V    A          C     A         U      F         A
\label{csmooth}

O algoritmo de gerenciamento de interesse proposto neste artigo, denominado \malg{} (ângulo de visão com área próxima e atenuação de freqüência de atualização), leva em conta três fatores principais:

\begin{itemize}
 \item Ângulo de visão do avatar, para determinar quais entidades o jogador tem que ser capaz de perceber imediatamente, por estarem à sua frente, até a distância que seu alcance de visão permita;
 \item Área próxima, cujo objetivo é %evitar que ocorram problemas caso o jogador faça seu avatar girar ao redor do próprio eixo muito rapidamente, além de 
 melhorar a qualidade do jogo no espaço mais perto do avatar. Seu raio é a distância crítica, definido anteriormente; %explicar melhor
 \item Atenuação da freqüência de atualizações.
\end{itemize}

A área de interesse resultante então toma a forma de um setor de círculo, cuja origem é o centro de outro círculo, menor. Este círculo menor é a área próxima do avatar do jogador, que receberá atualizações de estado com intervalo normal de entidades que nela estiverem. Dessa forma, tem-se o estado mais atualizado possível do jogo na região próxima ao avatar. Isso favorece a interação com entidades que estejam perto dele. Mesmo que alguma delas esteja momentaneamente fora do campo de visão do jogador, ela estará disponível caso ele gire seu avatar repentinamente na direção oposta à que está voltado. Na figura \ref{fcsmooth}, é ilustrada a área de interesse que acaba de ser definida.

Quanto às entidades que estiverem fora da área próxima, mas ainda dentro do ângulo de visão, será calculada sua relevância. Propõe-se que a relevância de cada entidade diminua gradualmente de acordo com a distância entre ela e o avatar do jogador em questão. Quanto mais longe, menos freqüentes serão as atualizações de estado. Isso é possível porque mesmo que o intervalo de atualização seja duplicado, ainda será de uma fração de segundo, o que será dificilmente perceptível por um jogador cujo avatar está situado a uma grande distância da entidade em questão. Além disso, pequenos atrasos entre a chegada das atualizações de estado podem ser facilmente mascarados através de técnicas de interpolação, como dead-reckoning \cite{smed2002rna}. O algoritmo \ref{smalg} define o funcionamento deste gerenciamento de interesse.

\begin{algorithm}
\caption{Calculate relevance of entity E to avatar A}
\label{smalg}
\begin{algorithmic}
 \STATE $dist \leftarrow distance(A, E)$
 \IF{$dist \le \cridis$}
 \STATE $relevance \leftarrow 1$
 \ELSE
 \IF{A can see E in its field of view}
 \STATE $relevance \leftarrow 1 - \frac{dist - \cridis}{\viewdis - \cridis}$
 \IF{$relevance < 0$}
 \STATE $relevance \leftarrow 0$
 \ENDIF
 \ELSE
 \STATE $relevance \leftarrow 0$
 \ENDIF
 \ENDIF
\end{algorithmic}
\end{algorithm}

\begin{figure}[!t]
	\centering
	%\includegraphics[width=2.5in]{max}
	\includegraphics[width=0.9\linewidth]{a3}
	\caption{\malg\ area of interest}
	\label{fcsmooth}
\end{figure}

\section{Simulation}
\label{sim}

Para efetuar a simulação do algoritmo proposto, foi necessário primeiro criar um modelo de ambiente virtual a simular, com diversos avatares presentes, pois o algoritmo é baseado nas informações de localização e ângulo de visão. O ambiente consiste em um espaço bidimensional, que corresponde à região gerenciada por um dos servidores. Nela, há diversos avatares presentes, cujo número varia de uma simulação para outra. Cada avatar escolhe aleatoriamente um ponto de destino no ambiente e segue até lá. Ao chegar no destino, permanece parado por um tempo aleatório, que pode ser zero, e então escolhe uma nova localização para se dirigir.

Para simular a técnica proposta, foi utilizado o simulador de rede ns-2 \cite{mccanne:nsn}. Este simulador permite criar código específico da aplicação que será simulada. No caso, foi simulado um servidor, que deveria enviar atualizações de estado para um cliente, responsável por um dos avatares na região. Baseado na localização dos outros avatares e no algoritmo de gerenciamento de interesse escolhido, o servidor decidia quais outros avatares tinham um estado relevante para o cliente em questão. Com isso, obtém-se a ocupação de largura de banda de envio necessária para um único cliente. Não se julgou necessário simular simultaneamente todos os clientes conectados àquele servidor, pois todos os avatares têm o mesmo perfil. Para encontrar a carga total no servidor, basta multiplicar a banda de envio necessária para um cliente pelo número de clientes presentes na região.

Outra questão é que o consumo de largura de banda de envio do servidor é muito maior que o de recepção - se ele recebe $n$ ações, cada uma oriunda de um dos $n$ clientes, é necessário, no pior caso, enviar $O(n^2)$ atualizações de estado, pois cada jogador precisaria do estado de todos os outros. Assim sendo, foi necessário apenas medir a banda de transmissão utilizada.

%falar do svoboda e do kim??...
Em trabalhos como \cite{yu2007nas}, \cite{kim2005tcm} e \cite{svoboda2007taa}, é analisado o tráfego de rede gerado por jogos em larga escala. Baseado nestes trabalhos, e adotando uma postura conservadora, foram decididos os seguinte parâmetros para serem utilizados na simulação:

\begin{itemize}
 \item Intervalo normal de atualização: 250 ms;
 \item Tamanho do pacote de atualização de estado de uma única entidade: 100 bytes;
 \item Duração de cada sessão de jogo simulada: 20 min;
 \item Área do ambiente virtual: 750 x 750;
 \item Alcance da visão: 120;
 \item Distância crítica: 40;
 \item Ângulo de visão: 180°.
\end{itemize}

Foram executadas diversas simulações, com o objetivo de comparar os algoritmos de gerenciamento de interesse apresentados. O número de avatares presentes no ambiente foi uma das variáves analisadas, para verificar a escalabilidade. Os algoritmos comparados foram os baseados em círculo, círculo com atenuação, ângulo de visão e o algoritmo proposto, \malg. Para demonstrar o quanto cada um destes reduz o tráfego, foram feitas simulações também em que não é empregado nenhum tipo de gerenciamento de interesse, e o servidor envia para o cliente atualizações de estado de todas as outras entidades do jogo.

\section{Results}
\label{results}

Os resultados foram coletados da seguinte maneira: para encontrar a largura de banda utilizada em média para envio, foram somados todos os pacotes de cada sessão e dividido pelo tempo que foi simulado; para determinar a largura de banda máxima utilizada, foi verificado, segundo a segundo, quantos bytes foram enviados e foi selecionado o máximo.

Nas tabelas \ref{tab_max} e \ref{tab_avg}, são apresentados os dados coletados de largura de banda máxima e média, respectivamente, utilizada com os quatro algoritmos simuladas - área em círculo (\circleaoi), área em círculo com atenuação (\circleat), área do campo de visão (\fov) e área do campo de visão mais área próxima mais atenuação (\malg) - além de mostrar quanto seria a largura de banda utilizada se nenhuma técnica fosse empregada (\noaoi). Os valores estão em bytes/s. Nas figuras \ref{fig_max} e \ref{fig_avg} são mostrados os gráficos correspondentes.

\begin{table}[ht]
\caption{Largura de banda máxima utilizada}
\centering
  \begin{tabular}{ c | c c c c c }  
    \hline
   	Avatars & \noaoi & \circleaoi & \circleat & \fov & \malg{} \\ \hline
		25	&	9400	&	8500	&	5700	&	7100	&	4700 \\
		50	&	19300	&	17000	&	10300	&	12300	&	8100 \\
		75	&	29100	&	23600	&	16600	&	17800	&	11300 \\
		100	&	38800	&	32500	&	20500	&	23000	&	15500 \\
		125	&	48600	&	37400	&	24300	&	29500	&	19700 \\
		150	&	58300	&	47400	&	29900	&	32900	&	22700 \\
		175	&	67700	&	56100	&	34300	&	32400	&	21500 \\
		200	&	77600	&	62300	&	37500	&	41200	&	28900 \\
	\hline
  \end{tabular}
\label{tab_max}
\end{table}

\begin{table}[ht]
\caption{Largura de banda utilizada em média}
\centering
  \begin{tabular}{ c | c c c c c }  
    \hline
   	Avatars & \noaoi & \circleaoi & \circleat & \fov & \malg{} \\ \hline
		25	&	9221	&	4715	&	2759	&	2534	&	1700 \\
		50	&	18826	&	9350	&	5442	&	4949	&	3303 \\
		75	&	28432	&	13963	&	8315	&	7619	&	5137 \\
		100	&	38037	&	19324	&	11029	&	9928	&	6739 \\
		125	&	47642	&	23138	&	13871	&	12434	&	8290 \\
		150	&	57247	&	29031	&	16432	&	15085	&	10062 \\
		175	&	66853	&	34697	&	19661	&	23060	&	14250 \\
		200	&	76458	& 	38600	&	23450	&	21491	&	14413 \\
	\hline
  \end{tabular}
\label{tab_avg}
\end{table}

\begin{figure}[!t]
	\centering
	%\includegraphics[width=2.5in]{max}
	\includegraphics[width=0.9\linewidth]{max}
	\caption{Simulation Results: maximum bandwidth usage}
	\label{fig_max}
\end{figure}

\begin{figure}[!t]
	\centering
  %\includegraphics[width=2.5in]{avg}
	\includegraphics[width=0.9\linewidth]{avg}
	\caption{Simulation Results: average bandwidth usage}
	\label{fig_avg}
\end{figure}

Apenas usando diferentes freqüências de atualização no gerenciamento de interesse baseado em círculo, reduziu-se em 41.59\% a largura de banda de envio utilizada em média pelo servidor por cliente. A utilização máxima de largura de banda também foi reduzida, em 36.19\%. Estes valores representam a média de redução de uso de largura de banda para os diferentes números de clientes.

No que diz respeito ao algoritmo proposto \malg, obteve-se uma redução de uso médio de largura de banda de envio de 63.51\% e 33.58\%, comparado respectivamente com o algoritmo de área de interesse circular e baseado em ângulo de visão. Reduziu-se também o pico de utilização em 52.03\% e 33.10\%, comparado com os mesmos algoritmos. Na tabela \ref{tab_summary}, são mostrados os percentuais médios de economia de largura de banda máxima e média com o algoritmo \malg, em relação aos outros algoritmos apresentados.

\begin{table}[ht]
\caption{Economia de largura de banda com o algoritmo \malg}
\centering
  \begin{tabular}{ c | c c c c }  
    \hline
   	Utilização & \noaoi & \circleaoi & \circleat & \fov \\ \hline
		Máxima	&	60.10\%	&	52.03\%	&	24.81\%	&	33.10\%	\\
		Média	&	81.64\%	&	63.51\%	&	37.48\%	&	33.58\%	\\
	\hline
  \end{tabular}
\label{tab_summary}
\end{table}

Observou-se também que os valores médio e máximo observados diferem, mesmo quando não é utilizado nenhum algoritmo de gerenciamento de interesse, ou seja, o cliente recebe atualizações de estado de todas as entidades presentes no jogo, com a freqüência normal. Além disso, com 200 avatares no ambiente, com estado de 100 bytes, cuja atualização é enviada a cada 250 ms, o servidor deveria alocar 199$\times$100$\times$4 bytes/s para cada cliente, ou seja, 79600 bytes/s. No entanto, observou-se que a utilização máxima e média, com 200 avatares presentes e nenhum gerenciamento de interesse, foi de 77600 e 76458, respectivamente. Isso acontece porque o ns-2 é um simulador de eventos discreto, e o servidor simulado foi programado para checar o schedule de envios a cada 10 ms. Em conseqüência disto, cada atualização de estado pode ter tido seu intervalo aumentado em até 10 ms, o que explica os valores encontrados.

\section{Conclusion}
\label{conclusion}
Foi apresentado um algoritmo de gerenciamento de interesse, o \malg, cuja idéia principal é adaptar a freqüência de atualização de estado das entidades do jogo de acordo com sua relevância para o cliente que receberá as atualizações. O formato da área de interesse utilizada pelo algoritmo \malg{} consiste em um setor de círculo, correspondente ao campo de visão do jogador, mais um círculo de raio menor, que corresponde à área próxima ao avatar daquele jogador. O objetivo deste círculo menor é o de manter o estado naquela região, que é considerada crítica, o mais atualizado possível. Somando-se essas características, chegamos a um algoritmo que obteve redução da utilização máxima da banda de envio do servidor de 52.03\% e 33.10\%, comparados com o gerenciamento de interesse baseado em círculo e em campo de visão, respectivamente, e de 63.51\% e 33.58\% de utilização média, comparados com os mesmos algoritmos.

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation Results}
%\label{fig_sim}
%\end{figure}

% Note that IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command, the
% \label for the overall figure must come after \caption.
% \hfil must be used as a separator to get equal spacing.
% The subfigure.sty package works much the same way, except \subfigure is
% used instead of \subfloat.
%
%\begin{figure*}[!t]
%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
%\label{fig_second_case}}}
%\caption{Simulation results}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.

% An example of a floating table. Note that, for IEEE style tables, the 
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that IEEE does not put floats in the very first column - or typically
% anywhere on the first page for that matter. Also, in-text middle ("here")
% positioning is not used. Most IEEE journals/conferences use top floats
% exclusively. Note that, LaTeX2e, unlike IEEE journals/conferences, places
% footnotes above bottom floats. This can be corrected via the \fnbelowfloat
% command of the stfloats package.

% conference papers do not normally have an appendix

% use section* for acknowledgement
%\section*{Acknowledgment}

%The authors would like to thank...

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
\bibliography{dsrt}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
% \begin{thebibliography}{1}
% 
% \bibitem{IEEEhowto:kopka}
% H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
%   0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
% 
% \end{thebibliography}

% that's all folks
\end{document}