### 2009-01-05 textos/dissertation/a3.tex

\chapter {A${^3}$: um algoritmo de otimização por área de interesse}