### 2009-01-09 trunk/textos/dissertation/conclusao.tex

%Quebrar em dois capitulos, se necessário.
\chapter {Considerações Finais e Trabalhos Futuros}

%por TUDO do plano de doutorado!!!