### 2011-05-23 disciplinas/usi_algsem/paxosTalk/paxosTalk.tex
19 pages paxosTalk

%% LaTeX Beamer presentation template (requires beamer package)
%% see http://latex-beamer.sourceforge.net/
%% idea contributed by H. Turgut Uyar
%% template based on a template by Till Tantau
%% this template is still evolving - it might differ in future releases!

\newcommand{\bval}{$B_{val}$}
\newcommand{\bqrm}{$B_{qrm}$}
\newcommand{\bvot}{$B_{vot}$}
\newcommand{\bbal}{$B_{bal}$}
\newcommand{\bvalm}{B_{val}}
\newcommand{\bqrmm}{B_{qrm}}
\newcommand{\bvotm}{B_{vot}}
\newcommand{\bbalm}{B_{bal}}
\newcommand{\vp}{$v_{p}$}
\newcommand{\vval}{$v_{val}$}
\newcommand{\vbal}{$v_{bal}$}
\newcommand{\votesbeta}{$\text{\textit{Votes}}(\beta)$}
\newcommand{\maxvotep}{\textit{MaxVote(b, p, $\beta$)}}
\newcommand{\maxvoteq}{\textit{MaxVote(b, Q, $\beta$)}}
\DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}
\DeclareMathAlphabet{\mathcalligra}{T1}{calligra}{m}{n}

\documentclass[10 pt]{beamer}

\mode<presentation>
{
\usetheme{Warsaw}
\usecolortheme{seagull}

%\usecolortheme{dolphin}
%\usecolortheme{seahorse}
%\usecolortheme{albatross}
\setbeamercovered{transparent}
}

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}

% font definitions, try \usepackage{ae} instead of the following
% three lines if you don't like this look

\usepackage{mathptmx}
%\usepackage[scaled=.90]{helvet}
\usepackage{courier}
\usepackage{chancery}
\usepackage{mathrsfs}
\usepackage{amsmath}
\usepackage{upgreek}

\usepackage[T1]{fontenc}

%\title{The Part-Time Parliament}
\title[{\makebox[.45\paperwidth]{The Part-Time Parliament\hfill\insertframenumber/\inserttotalframenumber}}]{The Part-Time Parliament} 
\subtitle{Leslie Lamport, 1998}

% - Use the \inst{?} command only if the authors have different
%   affiliation.
%\author{F.~Author\inst{1} \and S.~Another\inst{2}}
\author{Carlos Eduardo B. Bezerra}

% - Use the \inst command only if there are several affiliations.
% - Keep it simple, no one is interested in your street address.
\institute[Universities of]
{
Faculty of Informatics\\
Universittà della Svizzera italiana
}

\date{May 25th, 2011}

% This is only inserted into the PDF information catalog. Can be left
% out.
\subject{Talks}

% If you have a file called "university-logo-filename.xxx", where xxx
% is a graphic format that can be processed by latex or pdflatex,
% resp., then you can add a logo as follows:

% \pgfdeclareimage[height=0.5cm]{university-logo}{university-logo-filename}
% \logo{\pgfuseimage{university-logo}}

% % Delete this, if you do not want the table of contents to pop up at
% % the beginning of each subsection:
% %\AtBeginSubsection[]
% %{
% %\begin{frame}<beamer>
% %\frametitle{Outline}
% %\tableofcontents[currentsection,currentsubsection]
% % %\end{frame}
%}

% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

%\beamerdefaultoverlayspecification{<+->}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}
  \frametitle{Outline}
  \tableofcontents
  % You might wish to add the option [pausesections]
\end{frame}

%==========================================================
\section{Introduction}
%==========================================================

%----------------------------------------------------------
\subsection{A long way to Paxos}
%----------------------------------------------------------

\begin{frame}
  \frametitle{Introduction}

  Interesting information about `The Part-Time Parliament':

  \begin{itemize}
    \item the algorithm was first described in 1989 as a technical report [1]
    \item the paper was submitted in 1990\ldots and accepted in 1998%people just couldn't understand it!
    \item several other publications, such as [2, 3, 4], had explaining it as their main purpose\ldots
  \end{itemize}

\end{frame}

\begin{frame}

  The paper mostly uses an analogy with the fictional ancient island of Paxos, where:
  \begin{itemize}
    \item The parliament's (group's) legislators (processes) and their messengers (communication channels) may be unpredictably absent;
    \item There may be several ballots for each decree to be passed (rounds in each consensus instance);
    \item Legislators -- and messengers -- may be lazy or absent, but not dishonest (non-Byzantine failure model*)
  \end{itemize}
  \vspace{6 mm}
  {\small * The algorithm was extended later, however, to tolerate Byzantine failures [5]}
\end{frame}

%==========================================================
\subsection{The consensus problem}
%==========================================================

\begin{frame}
  \frametitle{Definition}

  \textbf{Consensus} is the process of agreeing on one single result among a group of participants. Assuming a collection of processes that can propose values, a consensus algorithm ensures that:

  \begin{itemize}
    \item only a single value among the proposed ones may be chosen;
    \item if no value has been proposed, then no value is chosen;
    \item a process never learns that a value has been chosen unless it really has been.
  \end{itemize}

%A process may be unaware of a given chosen value, as long as it doesn't learn an invalid one

%These properties guarantee consistency (safety), but not progress (liveness)
%Hard because of assuming the possibility of process failures\ldots

\end{frame}

%==========================================================
\section{The Paxos Algorithm}
%==========================================================

%----------------------------------------------------------
\subsection{Notations used}
%----------------------------------------------------------

\begin{frame}
  \frametitle{The Paxos algorithm}
  \framesubtitle{Some notations used}

  A value is chosen after a series of numbered \textbf{ballots}. For each ballot $B$, we have that:
  \begin{itemize}
    \item \bbal\ is its ballot number; %loosely related to its order
    \item \bqrm\ is a nonempty finite set of processes (quorum);
    \item \bvot\ is the set of processes who voted in $B$ -- each process may decide to participate (vote) or not in $B$;
    \item \bval\ is the value being voted on in that ballot.
  \end{itemize}

  \vspace{4 mm}
  A ballot is said to be \emph{successful} when \bqrm\ $\subseteq$ \bvot\ -- which means that every quorum member voted.

\end{frame}

\begin{frame}
  \frametitle{The Paxos algorithm}
  \framesubtitle{Some more notations}

  In each ballot, each process can cast a vote (accept or not a given proposed value). A vote $v$ has the components:

  \begin{itemize}
    \item \vp, which is the process $p$ who cast it;
    \item \vbal, the number of the ballot for which it was cast;
    \item \vval, a value voted on.
  \end{itemize}

  \vspace{4 mm}
 % A \emph{null} vote $v$ is that for which \vbal\ $=$ $-\infty$ and \vval\ $=$ BLANK. 
 For the set $\beta$ of ballots, the set \votesbeta\ is defined as\\
 $\{v$ : $\exists B$ $\in$ $\beta$ | \vp\ $\in$ \bvot\ $\wedge$ \vbal\ $=$ \bbal\ $\wedge$ \vval\ $=$ \bval\ $\}$

\end{frame}

\begin{frame}
  \frametitle{The Paxos algorithm}
  \framesubtitle{Still more notations\ldots}

  We then define \maxvotep\ as the vote $v$ with highest \vbal\ in 
 $\{v \in Votes(\beta) : (v_p = p) \wedge (v_{bal} < b)\} \cup \{null\}$,\\
 where $null$ means that $p$ didn't vote in any $B \in \beta$ ($null_{bal} = -\infty$)

  \vspace{4 mm}
  We also define \maxvoteq\ as the vote $v$, such that \vbal\ is the highest among all \maxvotep\ with $p \in Q$

  \vspace{8 mm}
  {\small With such notations, we can specify the conditions which ensure consistency and allow for progress.}

\end{frame}

%----------------------------------------------------------
\subsection{Consistency conditions}
%----------------------------------------------------------

\begin{frame}
  \frametitle{Consistency conditions}

  Consistency is guaranteed if the following conditions are ensured:

%write conditions on the whiteboard
  \begin{itemize}
    \item $B1(\beta)$

    $\forall B, B' \in \beta : (B \neq B') \Rightarrow (\bbalm \neq B'_{bal})$

    \vspace{4 mm}
    \item $B2(\beta)$ %B2 may be ensured by always picking a majority of processes as quorum, which implies that n > 2f (a majority is correct for each instance)

    $\forall B, B' \in \beta : \bqrmm \cap B'_{qrm} \neq \emptyset)$

    \vspace{4 mm}
    \item $B3(\beta)$

    $\forall B, B' \in \beta :$\\
    $(\text{\textit{MaxVote}}(\bbalm, \bqrmm, \beta)_{bal} \neq -\infty) \Rightarrow$\\
    $\Rightarrow (\bvalm = \text{\textit{MaxVote}}(\bbalm, \bqrmm, \beta)_{val})$
  \end{itemize}

\end{frame}

%----------------------------------------------------------
\subsection{Proof sketch}
%----------------------------------------------------------

\begin{frame}
\frametitle{Proof sketch}

%write lemma on the whiteboard
\textsc{\textbf{Lemma}}. If $B1(\beta)$, $B2(\beta)$ and $B3(\beta)$ hold, then $$((\bqrmm \subseteq \bvotm) \wedge (B'_{bal} > \bbalm)) \Rightarrow (B'_{val} = \bvalm)$$ for any $B$, $B'$ in $\beta$.

\vspace{3 mm}
\textsc{\textbf{Proof of Lemma}}. Let $\Psi(B, \beta)$ be defined as: $$\Psi(B, \beta) = \{B' \in \beta : (B'_{bal} > \bbalm) \wedge (B'_{val} \neq \bvalm)\}$$

Assume that $\bqrmm \subseteq \bvotm$. By contradiction, assume that $\Psi(B, \beta) \neq \emptyset$.

\vspace{3 mm}
(1) As $\Psi(B, \beta) \neq \emptyset$, we pick $C \in \Psi(B, \beta)$, such that $C_{bal} = min\{B'_{bal} : B' \in \Psi(B, \beta)\}$

\vspace{2 mm}
(2) From (1) and from the definition of $\Psi(B, \beta)$, we have that $C_{bal} > \bbalm$

\end{frame}

\begin{frame}
\frametitle{Proof sketch}
\framesubtitle{continuing\ldots}

(3) As $\bqrmm \subseteq \bvotm$ and $\forall B, B' \in \beta : \bqrmm \cap B'_{qrm} \neq \emptyset$, we have that $\bvotm \cap C_{qrm} \neq \emptyset$

\vspace{2 mm}
(4) As $C_{bal} > \bbalm$ and $\bvotm \cap C_{qrm} \neq \emptyset$, we have that \textit{MaxVote}$(C_{bal}, C_{qrm}, \beta)_{bal} \geq \bbalm$

\vspace{2 mm}
(5) From (4) and the definition of \textit{MaxVote}$(C_{bal}, C_{qrm}, \beta)$, we know that such vote exists in \textit{Votes}$(\beta)$ (i.e. it is not $null$).

\vspace{2 mm}
(6) From (5) and $B3(\beta)$, we have that \textit{MaxVote}$(C_{bal}, C_{qrm}, \beta)_{val} = C_{val}$

\vspace{2 mm}
(7) From (6), and as $C_{val} \neq \bvalm$ (from $\Psi(B, \beta$)), we have that \textit{MaxVote}$(C_{bal}, C_{qrm}, \beta)_{val} \neq \bvalm$

\end{frame}

\begin{frame}
\frametitle{Proof sketch}
\framesubtitle{continuing\ldots}

(8) As \textit{MaxVote}$(C_{bal}, C_{qrm}, \beta)_{bal} \geq \bbalm$ (from (4)) and \textit{MaxVote}$(C_{bal}, C_{qrm}, \beta)_{val} \neq \bvalm$ (from (7)), we have that \textit{MaxVote}$(C_{bal}, C_{qrm}, \beta)_{bal} > \bbalm$

\vspace{2 mm}
(9) From (8), and since \textit{MaxVote}$(C_{bal}, C_{qrm}, \beta)_{val} \neq \bvalm$, we have that \textit{MaxVote}$(C_{bal}, C_{qrm}, \beta) \in \text{\textit{Votes}}(\Psi(B,\beta))$

\vspace{2 mm}
(10) By the definition of \textit{MaxVote}$(C_{bal}, C_{qrm}, \beta)$, we have that \textit{MaxVote}$(C_{bal}, C_{qrm}, \beta)_{bal} < C_{bal}$

\vspace{2 mm}
(11) From (10), we have that $\exists B' : B' \in \Psi(B, \beta) \wedge B'_{bal} < C_{bal}$, which means that $$C_{bal} \neq min\{B'_{bal} : B' \in \Psi(B, \beta)\}$$ which is a contradiction with (1). $\square$

\end{frame}

\begin{frame}
\frametitle{Proof sketch}
\framesubtitle{continuing\ldots}

With the lemma, we show that, for a given set $\beta$ of ballots which hold $B1$, $B2$ and $B3$, then any two successful ballots decided the same value.

\vspace{2 mm}
\textbf{\textsc{Theorem 1}}. If $B1(\beta)$, $B2(\beta)$ and $B3(\beta)$ hold, then
$$((\bqrmm \subseteq \bvotm) \wedge (B'_{qrm} \subseteq B'_{vot})) \Rightarrow (B'_{val} = B_{val})$$
for any $B$, $B'$ in $\beta$.

\vspace{2 mm}
\textbf{\textsc{Proof of Theorem}}. If $B'_{bal} = B_{bal}$ then $B1(\beta)$ implies $B' = B$. If $B'_{bal} \neq B_{bal}$, the theorem follows immediately from the lemma. $\square$

\end{frame}

\begin{frame}
\frametitle{Proof sketch}
\framesubtitle{continuing\ldots}

If there are enough correct (non-faulty) processes, then a new ballot may be conducted while $B1$, $B2$ and $B3$ are preserved.

\vspace{2 mm}
\textbf{\textsc{Theorem 2}}. Let $b$ be a ballot number, and let $Q$ be a set of processes such that $b > \bbalm$ and $Q \cap \bqrmm \neq \emptyset$, for all $B \in \beta$. If $B1(\beta)$, $B2(\beta)$ and $B3(\beta)$ hold, then there is a ballot $B'$ with $B'_{bal} = b$ and $B'_{qrm} = B'_{vot} = Q$ such that $B1(\beta \cup \{B'\})$, $B2(\beta \cup \{B'\})$ and $B3(\beta \cup \{B'\})$ hold.

\vspace{2 mm}
\textbf{\textsc{Proof of Theorem}}.
\begin{itemize}
\item $B1(\beta \cup \{B'\})$ follows from $B1(\beta)$ and the fact that $B'_{bal} = b$, and the assumption about $b$.
\item $B2(\beta \cup \{B'\})$ follows from $B2(\beta)$ and the fact that $B'_{qrm} = Q$, and the assumption about $Q$.
\item For $B3(\beta \cup \{B'\})$, if \textit{MaxVote}$(b,Q,B) = -\infty$ then let $B'_{val}$ be any value; otherwise, let $B'_{val} = \text{\textit{MaxVote}}(b,Q,B)_{val}$. $B3(\beta \cup \{B'\})$ then follows from $B3(\beta)$. $\square$
\end{itemize} 
\end{frame}

%----------------------------------------------------------
\subsection{Preliminary protocol}
%----------------------------------------------------------

\begin{frame}
\frametitle{Preliminary protocol}

A protocol can be derived directly from $B1(\beta)$, $B2(\beta)$ and $B3(\beta)$, so that all these properties are maintained throughout its execution and a value may* be agreed upon.

\begin{itemize}
  \item To maintain $B1(\beta)$, whenever $p$ starts a new ballot $B$, it assigns a sequence number $seq$, which was never used by $p$. To avoid collisions, $p$'s unique id is appended. Then $\bbalm = (seq, p_{id})$.
  \item To maintain $B2(\beta)$, each ballot $B$ has \bqrm\ as the majority set of the processes. As it is impossible to choose two disjoint majority sets from the same group*, $B2(\beta)$ is satisfied.
  \item To maintain $B3(\beta)$, before initiating a ballot $B$, $p$ has to find out what is \textit{MaxVote}$(B, \bqrmm, \beta)$, which is done by finding the \textit{MaxVote}$(B, q, \beta)$ for each $q$ in \bqrm.
  \begin{itemize}
    \item If \textit{MaxVote}$(B, \bqrmm, \beta) = null$, then \bval\ can be anything;\\
    otherwise $\bvalm = \text{\textit{MaxVote}}(B, \bqrmm, \beta)_{val}$.
  \end{itemize}
\end{itemize}

{\small * If half of the processes is faulty, consistency is still respected and there is no deadlock, but the ballot $B$ does not succeed}

\end{frame}

\begin{frame}
\frametitle{Preliminary protocol}%all the phases are optional! consistency is always guaranteed!

  Phase 1) Find \textit{MaxVote}$(B, \bqrmm, \beta)_{val}$

  \hspace{3 mm} \textbf{Phase 1a}) $p$ picks a ballot number $b$ and sends \textit{NextBallot}$(b)$ to other processes

  \hspace{3 mm} \textbf{Phase 1b}) each process $q$ responds with \textit{LastVote}$(b, vote)$ to $p$, where $vote$ is the vote cast by $q$ with the highest ballot number less than $b$ (or $null$)

  \vspace{6 mm}

  Phase 2) Execute the ballot $B$

  \hspace{3 mm} \textbf{Phase 2a}) after receiving \textit{LastVote}$(b, vote)$ from a majority $Q$ of the processes, $p$ initiates the ballot with number $b$ and value $val$ such that $B3(\beta)$ is maintained, by sending \textit{BeginBallot}$(b, val)$ to every process in $Q$

  \hspace{3 mm} \textbf{Phase 2b}) upon receiving \textit{BeginBallot}$(b, val)$, a process will not vote on it if it has already sent \textit{LastVote}$(b', val')$, such that $b' > b$; otherwise, it sends \textit{Voted}$(b, q)$ to $p$ and records this vote in its memory

  \vspace{6 mm}

\end{frame}

\begin{frame}
\frametitle{Preliminary protocol}

  Phase 3) Announce the result of consensus

  \hspace{3 mm} \textbf{Phase 3a}) if $p$ receives \textit{Voted}$(b, q)$ from a majority $Q$ of processes, then it sends \textit{Success}$(val)$ to all processes

  \hspace{3 mm} \textbf{Phase 2b}) upon receiving \textit{Success}$(val)$, the process learns that the group achieved consensus on $val$.

\end{frame}

%==========================================================
\section*{Summary}
%==========================================================

\begin{frame}
\frametitle<presentation>{Summary}

\begin{itemize}
  \item The \alert{first main message} of your talk in one or two lines.
\end{itemize}

% The following outlook is optional.
\vskip0pt plus.5fill
\begin{itemize}
  \item Outlook
  \begin{itemize}
    \item Something you haven't solved.
    \item Something else you haven't solved.
  \end{itemize}
\end{itemize}

\end{frame}

%==========================================================
\section*{References}
%==========================================================

\begin{frame}
  \frametitle{References}

[1] Lamport, L. \textbf{The part-time parliament}, Technical Report 49, Systems Research Center, Digital Equipment Corp., 1989
\vspace{4 mm}

[2] Lamport, L. \textbf{Paxos Made Simple}, ACM SIGACT News, 2001
\vspace{4 mm}

[3] De Prisco, R., Lampson, B., Lynch, N. \textbf{Revisiting the Paxos Algorithm}, Distributed Algorithms, 1997
\vspace{4 mm}

[4] Lampson, B. \textbf{How to build a highly available system using consensus}, Distributed Algorithms, 1996
\vspace{4 mm}

[5] Castro, M., Liskov, B., \textbf{Practical Byzantine Fault Tolerance}, Operating Systems Design and Implementation, 1999

\end{frame}

\end{document}