### 2011-05-26 trunk/textos/optpaxos/draft.tex
quasi-genuine global fifo total order multicast pronto

\documentclass[times, 10pt]{article} 
%\documentclass[times, 10pt,twocolumn]{article} 
%\usepackage[noend]{distribalgo}
\usepackage{algorithm}
%\usepackage{times}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{color}
\usepackage{amssymb}
\usepackage{algorithm}
\usepackage[noend]{distribalgo}
\usepackage[draft]{fixme}

\begin{document}

\newcommand{\mv}[1]{\ensuremath{\operatorname{\mathit{#1}}}}
\definecolor{dark}{gray}{.6}
\newcommand{\bc}[1]{\textcolor{dark}{#1}}
\newtheorem{lems}{Lemma}
\newtheorem{props}{Proposition}
\newtheorem{thms}{Theorem}
\newtheorem{defs}{Definition}
\newtheorem{obs}{Observation}

\newcommand{\code}[1]{\texttt{\small{\textbf{#1}}}}

\newcommand{\blankline}[1]{\vspace{4 mm}}

\title{Practical consistency management for geographically distributed MMOG servers}

%\author{
%xxx\\
%University \\ Country\\
%\and
%xxx\\
%University \\ Country\\
%\and
% ...
%}

\maketitle

\begin{abstract}

\end{abstract}

\section{Introduction}
\label{sec:intro}

\section{System model and definitions}
\label{sec:model}

We assume a system composed of nodes, divided into \emph{players} and \emph{servers}, distributed over a geographical area. Nodes may fail by crashing and subsequently recover, but do not experience arbitrary behavior (i.e., no Byzantine failures). Communication is done by message passing, through the primitives \emph{send}$(p,m)$ and \emph{receive}$(m)$, where $p$ is the addressee of message $m$. Messages can be lost but not corrupted. If a message is repeatedly resubmitted to a \emph{correct node} (defined below), it is eventually received.

Our protocols ensure safety under both asynchronous and synchronous execution periods. The FLP impossibility result~\cite{FLP85} states that under asynchronous assumptions consensus cannot be both safe and live. We thus assume that the system is initially asynchronous and eventually becomes synchronous. The time when the system becomes synchronous is called the \emph{Global Stabilization Time (GST)}~\cite{DLS88}, and it is unknown to the nodes.
Before GST, there are no bounds on the time it takes for messages to be transmitted and actions to be executed. After GST, such bounds exist but are unknown. After GST nodes are either \emph{correct} or \emph{faulty}. A correct node is operational ``forever" and can reliably exchange messages with other correct nodes. This assumption is only needed to prove liveness properties about the system. In practice, ``forever" means long enough for one instance of consensus to terminate.

A game is composed of a set of objects. The game state is defined by the individual states of each one of its objects. We assume that the game objects are partitioned among different servers. Since objects have a location in the game, one way to perform this partitioning is by zoning the virtual world of the game. Each partition consists of a set of objects of the game and the server responsible for them is their \emph{coordinator}. As partitions represent physical regions in the game world, we define \emph{neighbor} regions as those which share a border with each other. We consider that each server is replicated, thus forming several groups which consist of the coordinator and its replicas. Therefore, for each region of the game world, there is a group assigned to it. This way, a group is said to coordinate an object when the group's coordinator is the coordinator for that object. Finally, groups are called neighbors when they are assigned to neighbor regions. From now on, the word `server' will be used for any kind of server, be it a coordinator or a replica.

Each player may send his command to one of the servers -- which might not be the group's coordinator, if that provides a lower delay between the issuing of a command and its delivery. In the case of avatar based virtual environments, and as an avatar is usually also an object, the server to which a player is connected belongs to the group of his avatar's coordinator. A command $C = \{ c_1, c_2, ... \}$ is composed of one or more subcommands, one subcommand per object it affects. We refer to the set of objects affected by command $C$ as $obj(C)$. Also, we refer to the objects coordinated by a server $S$ as $obj(S)$. Finally, we define $obj(C,S)$ = \{$o$ : $o \in obj(C)$ and $o \in obj(S)$\}.

%Each object has its own command stream. When a command affects more than one object, it must be inserted in the command stream of all the objects it affects.

Our consistency criterion is ``eventual linearizability". (I'm not sure this is indeed what we want and how to define it, but we do need some consistency criterion...)

%In order to guarantee that commands are executed properly by servers, theywe define the following order requirement.
% 
%Let $G =(V,E)$ be a graph where the set $V$ of vertices contains all game commands and $E$ contains edge $C_i \rightarrow C_j$ if subcommand $c_x \in C_i$ is executed before subcommand $c_y \in C_j$. The order requirement states that $G$ is acyclic. In other words, if two commands affect common objects, their subcommands must be executed in the same order.
%
%Our goal is to design ordering protocols with low latency. Two aspects affect latency: (a)~the number of communication steps needed to ensure that subcommands can be properly executed in the absence of contention and (b)~possible slowdowns due to contention. We illustrate these aspects in the following section.
%
%To each partition, a group of servers is assigned, and they are responsible for storing and managing the state of those objects. One of them is a coordinator (or region manager, if one considers partitions as regions) and the rest of the group is formed by replicas, whose purpose is to provide fault tolerance.

\section{Protocol}

To ensure reliability despite server crashes, each server is replicated using state-machine replication, implemented with Paxos. Each player sends his commands to the server to which he is connected, which then forwards it to the coordinator. Each command is assigned a timestamp and executed against objects in timestamp order. We implement this by using a logical clock in each server group. Guaranteeing that the same set of commands is executed by the respective affected objects in the same order provides the level of consistency we are seeking. Therefore, the challenge is how to assign timestamps to commands such that consistency is not violated and commands are not discarded due to stale timestamp values.

However, providing such level of consistency may prove to be costly in an MMOG context, since there may be several communication steps between the sending of a command and its atomic delivery. For this reason, we use a primitive we call quasi-genuine global fifo total order multicast, which is described in section \ref{sec:quasi}.

\subsection{Quasi-genuine Global FIFO Total Order Multicast}
\label{sec:quasi}

To reduce the time needed to deliver a message, we use a multicast primitive which delivers messages optmistically, based on the time when they are created in a sender. This \emph{optimistic} delivery, although not guaranteeing that all replicas will receive all messages, is designed to have a fairly low latency, counting from when each message is sent until it is delivered.

The final -- conservative, fault tolerant, but costly in terms of communication steps for each message -- delivery order should be as close as possible to the optimistic one, so that no rollbacks would be deemed necessary. To explain how this works, we first should understand the idea behind the optimistic delivery. Also, we must define what ``quasi-genuine'' and ``global FIFO'' means.

\textbf{Genuine} multicast protocols are those where two multicast groups only communicate with each other when one has some message to send to the other. A \textbf{quasi-genuine} multicast protocol assumes that:

\begin{center}
\emph{A1: every group knows from which other groups a multicast message can possibly arrive, and to which other groups it can send a multicast message.}
\end{center}

Different groups communicate with each other even if there is no message to be sent from one to the other, but, from A1, there is no need for a group $g_i$ to communicate with some other group $g_j$ which cannot send messages to $g_i$, or receive messages from $g_i$\footnote{This relation may even be asymmetric: it could be possible to send a message from group $g_i$ to $g_j$, but not in the opposite direction. In this case, in a protocol where a group is blocked waiting for possible messages from other groups, $g_j$ may block its delivery of messages waiting for some kind of ``clearance'' from $g_i$, but $g_i$ will never block waiting for messages from $g_j$.}. This information may be given by the application which is making use of such primitive, so, although not genuine, a quasi-genuine protocol does not imply that each group has to keep sending messages to every other group.

We define $sendersTo(G)$ as the set of groups which are able to send a message to $G$. Also, we define $receiversFrom(G)$ as the set of groups who are able to receive a message from $G$.

As messages are delivered according to their generation time, the delivery order is FIFO. Besides, as every two messages which are delivered in different groups should be delivered in the same order in these groups, we need total order. However, FIFO total order is not enough to guarantee one property that we need: let the send time $st(m)$ of a message $m$ be the value of the wallclock of $p$ when $m$ was sent, where $p$ is the process which sent it. We need that, if $st(m_1) < st(m_2)$, then $m_1$ should be delivered before $m_2$, if $m_1$ and $m_2$ are delivered by the same process. Although this property may be hard to observe, it is used by the optimistic delivery protocol, so that messages can be delivered in order in one communication step, in the best case. If, when each process sends a message $m$, it stamps the message with the current value $st(m)$ of its wallclock, we define:

\textbf{Global FIFO order}: if processes $p_i$ and $p_j$ send respectively $m_i$ and $m_j$ to the same destination $p_d$, then if $st(m_i) < st(m_j)$, $m_i$ is not delivered after $m_j$ in $p_d$.

Here, we consider that there are several processes. Each process $p$ belongs to a group $G = group(p)$, that is, $p \in G = group(p)$. Each group $G$ has a special leader process $c$, which is its coordinator.

\subsubsection{Optimistic delivery}

%Even when ordering commands relevant to only one group, a significant number of communication steps is required: player to replica; replica to coordinator; consensus (two communication steps in the case of Paxos, when having the coordinator as the leader, using ballot reservation, each acceptor sending phase 2b messages to all other acceptors and assuming a tolerable number of failures) and, once consensus is achieved, each replica can put the command in the delivery queue and send it to each player. That sums up to five communication steps. On a geographically distributed scenario, this number of communication steps may be prohibitive for the ``playability'' of a vast number of online games.

%In order to mitigate this problem, we propose a protocol for \emph{optimistic ordering and delivery} of the commands, which is supposed to run in parallel with the protocol described in the previous section, which from now on will be called \emph{conservative ordering and delivery}. The optimistic order and delivery are correct if they include all the messages delivered with the conservative one and if they are delivered in the same order. Also, each object has an optimistic state and a conservative state. When a command is conservatively delivered to an object, its conservative state is updated, and this state is certainly consistent -- given the assumptions we have made about the system. Likewise, when a command is optimistically delivered, its optimistic state is updated.

%The optimistic ordering and delivery is supposed to be much faster than the conservative one. Assuming that it will be, and if the optimistic delivery order is always correct -- which means that it correctly predicted the conservative delivery order --, then the objects will always have a valid state within much fewer communication steps, counting from when each command $C$ was issued to when it is delivered and applied to $obj(C)$.

The basic idea of the optimistic ordering is the following: assuming that the processes have a synchronized clock\footnote{We don't require here perfectly synchronized clocks, as the optimistic protocol tolerates mistakes by its very definition. We only need clocks which are synchronized enough, so that our delivery order prediction succeeds and matches the conservative delivery order.}, whenever a process $p$ (either coordinator or not) receives a message $m$ from a client, it immediately applies a timestamp $ts$ to it, which consists simply of the current value of $p$'s wallclock, $now$. Therefore, $m.ts=now$. A wait window of length $w(p)$ is considered, where $w(p)$ is defined as the highest value of the estimated communication delay plus the wallclock deviation between the process $p$ and any of the other processes in its group $G$ or in any of the groups in $sendersTo(G)$.

More formally, let $\delta(p',p)$ be the maximum time for a message from $p'$ to arrive at $p$. Also, let $\epsilon(p',p)$ be the deviation between the wallclocks of $p$ and $p'$. The value of $w(p)$ is estimated as the maximum value of $(\delta(p',p) + \epsilon(p',p))$ for every $p'$ in $G = group(p)$ or in $sendersTo(G)$.

After applying the timestamp to $m$, the process $p$ immediately forwards it to all the other processes involved, including those in other groups. A process is involved with a message $m$ when it is one of its destinations, which can be inferred from $m.dst$. Then, $p$ puts $m$ in an \mbox{\textit{optPending}} list, where it stays until $now>m.ts+w(p)$, which means that $m$ has been in that list for a time longer than the defined wait window. In the meantime, other messages, sent from other processes, may have been received and also inserted in that list, always sorting by their timestamps.

If $w(p)$ has been correctly estimated, and no message was lost, than all the messages that were supposed to be delivered before $m$ have necessarily been received already. If the same has occurred for all the processes, then all of them have received all the messages, and can deliver them in the same order of $ts$.

To avoid out-of-order deliveries, when a message $m$ arrives too late, that is, if when $m$ arrives at $p$, $m.ts < now - w(p)$, the message is simply discarded by $p$, since another message $m' : m'.ts > m.ts$ may have already been delivered\footnote{We could make this in a way such that $m$ is only discarded by $p$ if, in fact, there was a delivered message $m' : m'.ts > m.ts$. If there was no such message, $m$ could still be delivered without violating the order we defined.}.

However, even if the optimistic order is the same for all the processes, it won't be valid if the conservative order is different from it. For that reason, we devised a way to make the conservative delivery order as close as possible to the optimistic one.

\subsubsection{Conservative delivery}

Instead of having the conservative delivery done completely independently from the optimistic one, we can actually use the latter as a hint for the final delivery order. Since the optimistic delivery should be fast when compared to the conservative one, waiting for it should not decrease the system performance significantly. Also, if we wait a short period longer, we can avoid a rollback later caused by mismatches between the two delivery orders, which is not desirable. The basic idea is, then, to pick the optimistic delivery order seen at the coordinator of each group.

%The complicating factor is the possibility of a command affecting the state of objects in more than one group. So, all involved groups must somehow agree regarding the delivery order of these commands. However, as in most MMOGs objects interact only with other objects which are nearby, we can exploit the fact that the groups addressed by a single command are usually neighbors. We do this by defining \emph{barriers} for multicast, such that $barrier(G_{send},G_{recv})$ = $t$ means that the group $G_{send}$ promised that no commands with timestamp lower than $t$ would be sent to group $G_{recv}$ anymore. When a server $S$ has received all the barrier values from its neighbours, and they are all greater than a value $t$, then $S$ knows that no more commands are coming from other groups and that, once the local ordering is done, all the commands with timestamp up to $t$ can be conservatively delivered. Besides, a barrier is sent along with the bundle of all messages with timestamp greater than the last previous barrier sent from $G_{send}$ to $G_{recv}$, so that when a server has received a barrier from a group, it means that it knows all the messages sent by that group until the time value stored in that barrier.

The complicating factor is the possibility of a message having at least one destination group different from its source group. So, all involved groups must somehow agree regarding the delivery order of these messages. However, from assumption A1, each group knows which other groups it could send messages to -- or receive messages from. We can use this by defining \emph{barriers} for multicast, such that $barrier(G_{send},G_{recv})$ = $t$ means that the group $G_{send}$ promised that it would send no more messages with a timestamp lower than $t$ to group $G_{recv}$. We have defined that $sendersTo(G) =$ \mbox{$\{G' : G'\text{ is able to send a message to }G\}$}. When a process $p$, from group $G$, has received all the barrier values from all the groups in $sendersTo(G)$, and they are all greater than a value $t$, then $p$ knows that no more messages with timestamp lower than $t$ are coming from other groups and that, once the local ordering (the ordering of messages originated in $G$) is done, all the messages with timestamp up to $t$ can be conservatively delivered. Besides, a barrier is sent along with the bundle of all messages with timestamp greater than the last previous barrier sent from $G_{send}$ to $G_{recv}$, so that when a process has received a barrier from a group, it means that it knows all the messages sent by that group until the time value stored in that barrier.

There are three possibilities for each message $m$, in the perspective of the coordinator $c$ of a given group $G$:

\begin{itemize}
  \item The message $m$ was originated in $G$, which is the only destination of $m$:

  In this case, when $m$ is optimistically delivered by the coordinator $c$, $c$ also atomically broadcasts $m$ in $G$. When it is atomically delivered by a process $p$ of group $G$, it is not immediately delivered to the application. Instead, it is inserted in a \textit{consPending} list for later being conservatively delivered, once every group $G'$ in $sendersTo(G)$ have already sent a message $barrier(G',G) = t$, such that $t > m.ts$. This is done because there could be a message $m'$ yet to come from another group $G'$, such that $m'.ts < m.ts$.

  \item When $m$ is originated in $G$, but it has at least one group other then $G$ as a destination:

  In this case, when $m$ is optimistically delivered by the coordinator, it is also atomically broadcast, which is done by running a Paxos consensus instance within $G$, but having as learners the processes of the groups in $m.dst$. The proposed value is $\{m, barrier(G, G')\}$, where $barrier(G,G')$ is simply the value of $m.ts$. When the consensus is achieved, both the message and the barrier from $G$ are learnt by the processes of every destination group $G'$. The processes of all the involved groups also insert $m$ into the respective \textit{consPending} list.

  \item When $G$ is one of the destinations of $m$, but $m$ was originated in some other group $G'$:

  In this case, when $m$ is optimistically delivered by a process $p$ of group $G$, nothing else is done. The message $m$ should be inserted into the \mbox{\textit{consPending}} list of $p$ only when learnt via a consensus instance started at its group of origin -- stored in $m.src$.
\end{itemize}

The messages in the \textit{consPending} list are always sorted in ascending order of their timestamps. When the first message $m$, in the $consPending$ list of a process $p : group(p) = G$, is such that $m.ts < barrier(G, G')$ for all $G' \in sendersTo(G)$, then $m$ is conservatively delivered by $p$ to the application as the next message. We claim that this delivery respects the ``global FIFO total order''\footnote{Proof needed.}.

%The possibility not covered here is when a message is sent from a group to another one which is not its neighbour. However, it is realistic for this context to assume that objects should be close to each other in the virtual environment to interact and that, as such, this kind of interaction is not supported. Nevertheless, one simple way to provide this kind of support could be by each group waiting for the barriers of every other group in the system, instead of only those of its neighbors.

A more formal description of the protocol is given in Algorithm \ref{algorithm:quasi}. We consider that three primitives are given: $getTime()$, which returns the current value of the local wallclock; $Propose(val, learners)$, which initiates a Paxos \cite{paxosref} instance proposing the value $val$ with the local group as the group of acceptors and $learners$ as the group of learners; and also $Learn(val)$, which is called when a Paxos instance achieves consensus. The $Learn(val)$ is called for all the processes which are supposed to learn a given proposed value once it is decided by a Paxos instance. For the sake of simplicity, we assume that, for each group, the values are learnt by consensus in the same order in which the consensus instances were initiated\footnote{This can be easily done with a sequence number given by the group's coordinator when initiating each consensus instance -- but note that we don't care if consensus instances initiated by different groups are finished in a non-deterministic order, but only those initiated in each single group; we want something like a ``FIFO learn order'', and not necessarily atomic.}.

Moreover, each process $p$ of group $G$ keeps an \textit{optPending} list which contains the messages waiting to be \textit{OPT-Delivered} (optimistically delivered). The \mbox{\textit{consPending}} list contains the messages ready to be \textit{CONS-Delivered} (conservatively delivered), but which may be waiting for the clearance from the groups in $sendersTo(G)$. Also, each message $m$ has at least three fields: $dst$, which is the set of $m$'s destination groups; $src$, which is the group where $m$ was generated; and $ts$, the value of the wallclock of the process when $m$ was generated. Note that, by abuse of notation, we have `process $p \in m.dst$' instead of `process $p$ : $\exists$ group $G \in m.dst \wedge p \in G$'.

Something that must be noticed is that some messages might not have the source group as a destination. Because of that, to ensure that messages of this kind are also received by $group(p)$'s coordinator $c$, which then initiates the Paxos instance that allows for the conservative delivery of the message in all of its destinations, every message $m$ from $p$ must also be sent to $c$ (l. \ref{algline:sendcoord}). This way, when the clock value at the coordinator $time(c) > w(c) + m.ts$, $c$ will conservatively send $m$ to all the processes in $m.dst$. Finally, to avoid delivering a message $m$ in a group it is not addressed to, before OPT-Delivering it, each process must check whether it is one of $m$'s destinations or not (l. \ref{algline:notinterested}). In l. \ref{algline:sendcoord}, $m$ is sent to the whole group $G$, so that in the case of failure of $G$'s coordinator, it is more likely that the new coordinator will know about the existence of $m$.

\begin{algorithm}
\begin{distribalgo}[1]
%\blankline\
%\STATE\underline{This algorithm is executed by every process $p$, which belongs to group $G$:}
%\STATE
\blankline\
\INDENT {Initialization}
  \INDENT{\textbf{for all} $G' : G' \in sendersTo(G)$ \textbf{do}}
    \STATE $barrier(G',G) = -\infty$
  \ENDINDENT
\ENDINDENT
\blankline\
\INDENT{\textit{When $p$ generates a message $m$}}
  \STATE $m.ts \leftarrow getTime()$  
  \COMMENT{current wallclock value as the timestamp of $m$}
  \INDENT{\textbf{for all} $p' \in m.dst \cup G$ \textbf{do}} \label{algline:sendcoord}
    \STATE $send(p', m)$    
    \COMMENT{send optimistically $m$ to all involved processes}
  \ENDINDENT
\ENDINDENT
\blankline\
\INDENT{\textit{When a process $p$ receives a message $m'$}}
  \IF {$m'.ts < getTime() - w(p)$}
    \STATE discard $m'$
    \COMMENT{late commands probably lead to out-of-order delivery}
  \ELSE
    \STATE $optPending \leftarrow optPending$ $\cup$ $\{m'\}$
  \ENDIF
\ENDINDENT
\blankline\
\INDENT{\textit{When $\exists m \in optPending : getTime() > m.ts + w(p)$ $\wedge$\\$\nexists m' \in optPending: m'.ts < m.ts$}}
  \STATE $optPending \leftarrow optPending \setminus \{m\}$
  \IF {$p$ is a coordinator $\wedge$ $m.src = G$}
    \STATE $Propose(m, m.dst \cup G)$ \label{algline:propg}
  \ENDIF
  \IF {$G \in m.dst$}  \label{algline:notinterested}
    \STATE OPT-Deliver($m$)  
%   \INDENT{\textbf{for all} $cli \in clients(s)$ \textbf{do}}
%     \STATE $send(cli, \{c, \text{`optimistic'}\})$    
%     \COMMENT{s sends $c$ to its clients, warning it is optimistic}
%   \ENDINDENT
  \ENDIF  
\ENDINDENT
\blankline\
\INDENT{\textit{When} $Learn(m)$}
  \IF {$G \in m.dst$} \label{algline:checkcons}
    \STATE $consPending \leftarrow consPending \cup \{m\}$
  \ENDIF
  \IF {$m.src \neq G$}
    \STATE $barrier(m.src, G) = m.ts$
  \ENDIF
%   \INDENT{\textbf{for all} $cli \in clients(s)$ \textbf{do}}
%     \STATE $send(cli, \{c, \text{`final'}\})$    
%     \COMMENT{s sends $c$ to its clients, telling it is the final order}
%   \ENDINDENT
%   \IF {$p$ is a coordinator}
%     \STATE $consPending \leftarrow consPending \setminus \{c\}$
%   \ENDIF  
\ENDINDENT
\blankline\
\INDENT{\textit{When $\exists m \in consPending : \forall G' \in sendersTo(G): m.ts < barrier(G',G)$ $\wedge$ \\ $\nexists m' \in consPending : m'.ts < m.ts$ }}
  \STATE CONS-Deliver($m$)

\blankline\
\ENDINDENT

\caption{Delivery algorithm, executed by every process $p$ from group $G$}
\label{algorithm:quasi}
\end{distribalgo}
\end{algorithm}

\subsubsection{Addressing liveness}

The problem with Algorithm \ref{algorithm:quasi} is that it does not guarantee liveness when a group has no message to receive from some other group and then keeps waiting for a new message to increase the barrier value and proceed with the conservative delivery. However, it does not disrupt the optimistic delivery. Also, liveness for the conservative delivery can be easily provided by sending periodic empty messages from $G$ to each $G' \in receiversFrom(G)$ to which no message has been sent for a specified time threshold \textit{barrierThreshold}. Algorithm \ref{algorithm:liveness} describes this. For the coordinator of $G$ to know when a message was successfully delivered to other group via consensus, it must also learn it. For this reason, when it proposes any message for conservative delivery, it always includes $G$ as one of the group of learners (l. \ref{algline:propg} of Algorithm \ref{algorithm:quasi}). To avoid the conservative delivery of messages whose destinations do not include $G$, it checks before including it in the conservative delivery queue (l. \ref{algline:checkcons} of Algorithm \ref{algorithm:quasi}). This check also avoids delivering empty messages sent by other groups with the sole purpose of providing liveness.

\begin{algorithm}
\begin{distribalgo}[1]
\blankline\
\INDENT {Initialization}
  \INDENT{\textbf{for all} $G' : G' \in receiversFrom(G)$ \textbf{do}}
    \STATE $lastBarrierSent(G') = -\infty$
  \ENDINDENT
\ENDINDENT
\blankline\
\INDENT{\textit{When} $Learn(\{m\})$ $\wedge$ $m.src = G$}
  \INDENT{\textbf{for all} $G' : G' \in m.dst$ \textbf{do}}
    \STATE $lastBarrierSent(G') \leftarrow m.ts$
  \ENDINDENT
\ENDINDENT
\blankline\
\INDENT{\textit{When $\exists G' \in receiversFrom(G):$\\$getTime() - lastBarrierSent(G') > barrierThreshold$}}
  \STATE $null \leftarrow$ empty message
  \STATE $null.ts \leftarrow getTime()$
  \STATE $null.src \leftarrow G$
  \STATE $Propose(null, \{G,G'\})$
  \COMMENT{saving that nothing was sent until em.ts}
\ENDINDENT
\blankline\

\caption{Achieving liveness; executed by the coordinator $c$ of group $G$}
\label{algorithm:liveness}
\end{distribalgo}
\end{algorithm}

The problem with addressing liveness this way is that, in the worst case, $G'$ has decided a message $m$ and has just received the last barrier $b$ from $G$, such that $b$ is less than $m.ts$ by an infinitesimal difference. This would mean that, if $G$ has no messages to send to $G'$, $G'$ will have to wait for $barrierThreshold + 2\Delta$, assuming that $2\Delta$ is the time needed for a proposal to be learnt after it has been initiated. Because of that, the conservative delivery of $m$ will be delayed by $w(c') + 2\Delta + barrierThreshold + 2\Delta$, where $w(c')$ is the wait window of the coordinator of $G'$, and $2\Delta$ is counted twice because there are two consensus proposals -- first in $G'$ regarding $m$, and then in $G$ regarding the empty message.

There is a way to provide liveness withouth such a possibly big delay, although creating more messages and making deeper changes in the delivery algorithm. Let $blockers(m)$ be defined as the set of groups whose barrier is needed in order to some group to deliver $m$. More formally $blockers(m) = \{G_B : \exists G_{dst} \in m.dst \wedge G_B \in sendersTo(G_{dst})\}$. The idea is that, once each group $G_B$ in $blockers(m)$ have sent a barrier $b > m.ts$ to all the groups in $m.dst \cap receiversFrom(G_B)$, all possible destinations of $m$ can deliver it. This way, instead of relying on periodic messages, whenever a process $p$ in a group $G$ has a message $m$ to send -- no matter who are the destinations --, $p$ sends $m$ to every $p' \in m.dst \cup G \cup blockers(m)$. Like in Algorithm \ref{algorithm:quasi}, the message must be addressed to $G$, so that $G$'s coordinator starts a consensus instance to deliver it conservatively; and it is sent to the groups in $blockers(m)$, so that they know that there is a message which will be blocked until they send a proper barrier to unblock it.

When the coordinator $c'$ of some group $G'$ receives $m$, such that $m.src \neq G'$, $c'$ knows that there might be other groups depeding on $G's$ barrier to deliver $m$. For that reason, it will immediately create a $null$ message with a timestamp equals to $m.ts$ and insert it into its optimistic queue. As soon as $null.ts > now - w(c')$, the $null$ message will be proposed by $c'$ in $G'$, having \mbox{$m.dst \cap receiversFrom(G')$} as its learners -- this way, any group which was waiting for a barrier from $G'$ to deliver $m$ will be able to do so as soon as it learns the $null$ message which was just created. The new delivery algorithm would be as described in Algorithm {\ref{algorithm:ondemand}}.

\begin{algorithm}
\begin{distribalgo}[1]
%\blankline\
%\STATE\underline{This algorithm is executed by every process $p$, which belongs to group $G$:}
%\STATE
\blankline\
\INDENT {Initialization}
  \INDENT{\textbf{for all} $G' : G' \in sendersTo(G)$ \textbf{do}}
    \STATE $barrier(G',G) = -\infty$
  \ENDINDENT
\ENDINDENT
\blankline\
\INDENT{\textit{When $p$ generates a message $m$}}
  \STATE $m.ts \leftarrow getTime()$  
  \COMMENT{current wallclock value as the timestamp of $m$}
  \INDENT{\textbf{for all} $p' \in m.dst \cup G \cup blockers(m)$ \textbf{do}} \label{algline:askclearance}
    \STATE $send(p', m)$
  \ENDINDENT
\ENDINDENT
\blankline\
\INDENT{\textit{When a process $p$ receives a message $m'$}}
  \IF {$G \neq m'.src$} \label{algline:askedforclearance}
    \STATE $null \leftarrow$ empty message
    \STATE $null.src \leftarrow G$
    \STATE $null.ts \leftarrow m'.ts$
    \STATE $null.dst \leftarrow m'.dst \cap receiversFrom(G)$
    \STATE $optPending \leftarrow optPending \cup null$
  \ELSIF {$m'.ts < getTime() - w(p)$}
    \STATE discard $m'$
    \COMMENT{late commands probably lead to out-of-order delivery}
  \ELSE
    \STATE $optPending \leftarrow optPending$ $\cup$ $\{m'\}$
  \ENDIF
\ENDINDENT
\blankline\
\INDENT{\textit{When $\exists m \in optPending : getTime() > m.ts + w(p)$ $\wedge$\\$\nexists m' \in optPending: m'.ts < m.ts$}}
  \STATE $optPending \leftarrow optPending \setminus \{m\}$
  \IF {$p$ is a coordinator $\wedge$ $m.src = G$}
    \STATE $Propose(m, m.dst)$ %\label{algline:propg}
  \ENDIF
  \IF {$m \neq null$}  %\label{algline:notinterested}
    \STATE OPT-Deliver($m$)  
%   \INDENT{\textbf{for all} $cli \in clients(s)$ \textbf{do}}
%     \STATE $send(cli, \{c, \text{`optimistic'}\})$    
%     \COMMENT{s sends $c$ to its clients, warning it is optimistic}
%   \ENDINDENT
  \ENDIF  
\ENDINDENT
\blankline\
\INDENT{\textit{When} $Learn(m)$}
  \IF {$m \neq null$} %\label{algline:checkcons}
    \STATE $consPending \leftarrow consPending \cup \{m\}$
  \ENDIF
  \IF {$m.src \neq G$}
    \STATE $barrier(m.src, G) = m.ts$
  \ENDIF
%   \INDENT{\textbf{for all} $cli \in clients(s)$ \textbf{do}}
%     \STATE $send(cli, \{c, \text{`final'}\})$    
%     \COMMENT{s sends $c$ to its clients, telling it is the final order}
%   \ENDINDENT
%   \IF {$p$ is a coordinator}
%     \STATE $consPending \leftarrow consPending \setminus \{c\}$
%   \ENDIF  
\ENDINDENT
\blankline\
\INDENT{\textit{When $\exists m \in consPending : \forall G' \in sendersTo(G): m.ts < barrier(G',G)$ $\wedge$ \\ $\nexists m' \in consPending : m'.ts < m.ts$ }}
  \STATE CONS-Deliver($m$)

\blankline\
\ENDINDENT

\caption{Delivery algorithm, executed by every process $p$ from group $G$}
\label{algorithm:ondemand}
\end{distribalgo}
\end{algorithm}

Now, assuming that $w(c)$ is the same for every group coordinator $c$, we have a worst-case delivery latency for $m$ of $w(c) + 2\Delta$, as the $null$ message is created and proposed in parallel with $m$. However, each sender process is sending $m$ to every destination process -- which will not be changed --, but also to every process in its group and to every process belonging to some group of $blockers(m)$. This can create an amount of messages too high to be practical. A solution for that would be the following: instead of sending $m$ to every single process in $G$ and in each group of $blockers(m)$, the sender $p$ sends $m$ only to its destinations and to the coordinators of $G$ and of the groups in $blockers(m)$. That way, the line \ref{algline:askclearance} of Algorithm \ref{algorithm:ondemand} would be:
\\
\\
\textbf{for all}\\$p' \in m.dst \cup \{c : c\text{ coordinates }G \vee (\exists G_B \in blockers(m) \wedge c\text{ coordinates }G_B)\}$\\\textbf{do}

% Nevertheless, one possible way to implement the delivery for a command $C$ of this type would be to initiate a Paxos consensus instance in each group addressed by it, proposing $C$ and having as learns all servers in all involved groups.

%In brief, our strategy is to combine state-machine replication (implemented with Paxos), used to handle object replicas, with Skeen's algorithm, used to select timestamps for commands that affect multiple objects.

\subsection{Recovering from mistakes}

Unfortunately, even with a very good delay estimation (e.g. on an environment with a low jitter), there is absolutely no guarantee that the multicast protocol described  in section \ref{sec:quasi} will deliver the game command messages optimistically and conservatively in the same order. When it doesn't, it is considered a \emph{mistake}. Every mistake of the optimistic delivery -- either a lost command message, or an out-of-order delivery -- will cause a rollback of the optimistic state of the objects and re-execution of some of the optimistically delivered commands.

To perform that, we consider that each object has an optimistic delivery queue, $Q_{opt}$. Whenever a command is optimistically delivered, the optimistic state is updated and the command is pushed in the back of $Q_{opt}$. Whenever a command $C_c$ is conservatively delivered, it updates the conservative state of each object in $obj(C_c)$ and, for each one of them, the algorithm checks whether it is the first command in $Q_{opt}$. If it is, $C_c$ is simply removed from $Q_{opt}$ and the execution continues. If it isn't, it means that $C_c$ was either optimistically delivered out of order, or it was simply never optmistically delivered. It then checks whether $Q_{opt}$ contains $C_c$. If it does, it means the command was optimistically deliverd out of order, and it is removed from the list -- if $Q_{opt}$ doesn't contain $C_c$, it was probably lost\footnote{Also, when $C_c$ is delivered, but it is not in $Q_{opt}$, the remaining possibility is the very unlikely case where the conservative delivery happened before the optimistic one. To handle this case, $C_c$ is stored in a list of possibly delayed optimistic delivery and, if it is ever optimistically delivered, the algorithm will know that it should only discard that command, instead of updating the optimistic state.}. Then, the optimistic state is overwritten with the conservative one and, from that state, all the remaining comands in $Q_{opt}$ are re-executed, leading to a new optimistic state for that object.

%\bibliographystyle{latex8}
%\bibliography{main}

\end{document}